---
title: "Post Covid-19 Recovery"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
authors: Mariah Bastos Taytumn Schlegel Ridgway
---

Post COVID-19 Recovery 

# Project abstract
On 31 December 2019, WHO was alerted to several cases of pneumonia in Wuhan City, Hubei Province of China. The virus did not match any other known virus and has changed the world as we know it. The Novel Corona Virus forced entire countries to shut down with countries like Italy, Spain and France being hit exceptionally hard. An economic catastrophe shortly followed. In this analysis, we take a look at different sectors of the European economy to find key indicators of what may be expected in the Post-Covid world.  

# Problem statement 
Economies world-wide are experiencing extreme downturns with some of the worst coming from countries in Europe. It is imperative that with new data coming in everyday in a rapidly evolving environment, that we analyze the economy before Covid-19 as well as the economic change that has occurred during Covid-19 in order to make an accurate prediction of what can be expected in Post-Covid recovery. 

# Description
Taking data from Eurostat, a European statistics partnership with National Statistical Institutes and other national authorities in the EU Member States, known as the European Statistical System (ESS), we preform a variety of regression analyses along with a time series analysis on different key components of the European economy in order to make an accurate prediction of what may be expected in the European Post-Covid 19 economic recovery. 


# Load and Install the packages
```{r}
install.packages("tidyverse")
install.packages("caret")
install.packages("caTools")
install.packages("lubridate")
install.packages("dplyr")
install.packages("magrittr")
install.packages("ggpubr")
```
```{r}
library(magrittr)
library(tidyverse)
library(caTools)
library(lubridate)
library(dplyr)
library(ggpubr)
```

# Preparing the Data
## Load and Select the Data

```{r}
#Loading daily cases
numberOfCases <- read.csv('dataset/numberofcases.csv' )
ncdf <- data.frame(numberOfCases)

#Cases in Europe per month
ncdfEurope <- ncdf %>% 
  filter(continentExp == "Europe") %>%
  group_by(month, year, geoId) %>%
    summarize(TotalCases = sum(cases), TotalDeaths = sum(deaths)) %>%
    arrange(geoId,year,month)

#Cumulative Cases 
ncdfEuropeCumulative <- ncdfEurope %>% 
  group_by(geoId) %>%
  mutate("cumulativeDeaths" = cumsum(TotalDeaths)) %>%
  mutate("cumulativeCases" = cumsum(TotalCases)) %>%
  arrange(geoId,year,month)


#Interest Rate
interest <- read.delim('dataset/InterestByMonth.csv',sep =';')

#Creating dataframe
idf <- data.frame(interest)

#Joining the dataframe with cases in Europe
interestCases  <- right_join(ncdfEuropeCumulative,interest, by = c("geoId" = "Country", "month"="Month", "year" = "Year"))
#NA to 0
interestCases$TotalCases[is.na(interestCases$TotalCases)] <- 0
interestCases$TotalDeaths[is.na(interestCases$TotalDeaths)] <- 0
interestCases$cumulativeDeaths[is.na(interestCases$cumulativeDeaths)] <- 0
interestCases$cumulativeCases[is.na(interestCases$cumulativeCases)] <- 0

interestCasesHU <- interestCases %>%
  filter(geoId == 'HU') 

scatter.smooth(x=interestCasesHU$cumulativeCases, y=interestCasesHU$InterestRate, main="Interest Rates x Deaths - Hungary")


interestCasesHU.ts = ts(interestCasesHU$InterestRate, start = c(2019, 10), end = c(2020, 9), frequency = 5)

```
#Time Series

```{r}
#Time Series Hungary
summary(interestCasesHU.ts)

interestCasesBG <- interestCases %>%
  filter(geoId == 'BG') %>%
  arrange(geoId,year,month)

interestCasesBG.ts = ts(interestCasesBG$InterestRate, start = c(2019, 10), end = c(2020, 9), frequency = 5)

#Time Series Belgium
summary(interestCasesBG.ts)

interestCasesCZ <- interestCases %>%
  filter(geoId == 'CZ') %>%
    arrange(geoId,year,month)

scatter.smooth(x=interestCasesCZ$cumulativeCases, y=interestCasesCZ$InterestRate, main="Interest Rates x Deaths - 
Czech Republic")

interestCasesCZ.ts = ts(interestCasesCZ$InterestRate, start = c(2019, 10), end = c(2020, 9), frequency = 5)
#Time Series Czech Republic
summary(interestCasesCZ.ts)
```

# Variables
># Our independent variable: 
   - Number of cases
   Our dependent variables: 
   - Interest Rate

# Hypothesis Testing

## Hypothesis

># Null Hypothesis: There is no correlation between Covid-19 cases and key negative economic indicators in Europe
># Alternative Hypothesis: Covid-19 is correlated directly with key negative economic indicators in Europe.

## Visualisation

```{r}
ggdensity(interestCasesHU$InterestRate, fill = "lightgray",title = "Density - Interest Rate - Hungary"  )
ggqqplot(interestCasesHU$InterestRate, title =  "Q-Q - Interest Rate - Hungary")
ggdensity(interestCasesBG$InterestRate, fill = "lightgray",title = "Density - Interest Rate - Belgium"  )
ggqqplot(interestCasesBG$InterestRate, title =  "Q-Q - Interest Rate - Belgium")
```

# Normality Test

## Shapiro- Wilk Test

```{r}
shapiro.test(interestCases$TotalCases)
shapiro.test(interestCases$TotalDeaths)
shapiro.test(interestCases$cumulativeDeaths)
shapiro.test(interestCases$cumulativeCases)
```
># The data set is not normally distributed since p-value is less than alpha, therefore the distribution is significantly different from a normal distribution and we cannot assume normality. 

## Wilcox Test

```{r}
wilcox.test(interestCases$cumulativeCases,interestCases$InterestRate, PAIRED = FALSE)
```
># P value is smaller than alpha, we reject the null hypothesis. 


# Mean and standard deviation
 
```{r}
average<- interestCases %>%
            group_by(geoId) %>%
            summarise(average=mean(interestCases$InterestRate))
sdre=sd(interestCases$InterestRate, na.rm = TRUE)
sdre
```

# ANOVA Test

## Hypothesis

># Null Hypothesis: H0: m1 = m2 = m3

># Alternative Hypothesis: H1: m1  ≠ m2  ≠ m3

## Create Data Frame 

```{r}
anovaBG <- interestCases %>% 
  filter(geoId == "BG") 
anovaHU <- interestCases %>% 
  filter(geoId == "HU") 
anova <- interestCases 
```

## ANOVA

```{r}
aovResult <- aov(formula =  cumulativeCases~InterestRate , data = anova)
summary(aovResult)
aovResultBG <- aov(formula =  cumulativeCases~InterestRate , data = anovaBG)
summary(aovResultBG)
aovResultHU <-aov(formula =  cumulativeCases~InterestRate , data = anovaHU)
summary(aovResultHU)
```

># P-Value is lower than alpha, therefore we reject the null hypothesis that at all the means are equal and we conclude that at least one of the population means is significantly different. 

# Correlation

```{r}
Correlation <- interestCasesBG %>% ungroup() %>% select( c ("cumulativeDeaths", "InterestRate"))
pairs(Correlation)
```

## Correlation Calculation

```{r}
calculationcorBG <- cor(interestCasesBG$cumulativeCases, interestCasesBG$InterestRate,  use = "complete.obs")
calculationcorBG
calculationcorHU <- cor(interestCasesHU$cumulativeCases, interestCasesHU$InterestRate,  use = "complete.obs")
calculationcorHU
calculationcor <- cor(interestCases$cumulativeCases, interestCases$InterestRate,  use = "complete.obs")
calculationcor
```
># There is a weak positive correlation between Interest Rate and Cumulative Cases of COVID19, as the correlation is 0.3.

# Univariate Linear Regression
```{r}
model <- lm( cumulativeCases~InterestRate , data = interestCasesBG)
model
```

# Prediction
Using one country to test using the model.
```{r}
test <- interestCases %>%
          filter(geoId == "HU") 
            
predictedPosition <- predict(model, test)
predictedPosition

# Actual vs Predicted
# Create data frame with actual and predicted values
Prediction <- data.frame( predictedPosition,
                                test$TotalCases)
Prediction[is.na(Prediction)] <- 0
test$InterestRate[is.na(test$InterestRate)] <- 0
```

# Error reporting (mae, mse, rmse)

## MAE
```{r}
mae <- MAE(Prediction$predictedPosition, test$InterestRate)
cat("Mean Absolute Error:", mae)
```

## MSE
```{r}
MSE = function(m, o){
  (m - o)^2
}
mse <- MSE(Prediction$predictedPosition, test$InterestRate)
cat("Mean Squared Error", mse)
```

## RMSE
```{r}
RSME = function(m, o){
  sqrt(mean((m - o)^2))
}

rsme <- RSME(Prediction$predictedPosition, test$InterestRate)
cat("Mean Squared Error", rsme)
```

